Notes for ISIPTA poster

### 2016-07-12 asce-paper.tex
inv gamma with params a and b
rul paper, start of sec 2

Often, there is no failure data for the system itself
(e.g., if the system is a prototype, or the system is used under unique circumstances),
but there is some information about failure times for the components the system is made of.
We are able to analyse systems of arbitrary system structures, %layout of components,
i.e., any combination and nesting of series, parallel, k-out-of-n, or bridge-type arrangements,
by use of the survival signature \cite{2012:survsign}.
In this paper, we assume that components can be divided into $K$ different groups,
and components within each group $k$ ($k=1, \ldots, K$) can be assumed to follow the same failure time distribution.
We denote components of group $k$ as \emph{type $k$ components},
and assume independence between component types,
such that failure times of components of different types are independent.
We assume that type $k$ component lifetimes $T_i^k$ ($i = 1, \ldots, n_k$)
are Weibull distributed, with type-specific parameters $\beta_k$ (shape) and $\lambda_k$ (scale).
***type-specific shape not implemented yet!***

Focusing on the prototype application,
we assume that our observations consist solely of the failure times of components in this system up to time $\tnow$,
such that the failure times of components that have not failed by $\tnow$ are right-censored,
and calculate $\Rsys(t \mid t > \tnow) = P(\Tsys \geq t \mid t > \tnow)$,
which can be used to determine the remaining useful life (in short RUL, see, e.g., ***) of the system.
%This sparse information is complemented by expert information on component lifetimes $T_i^k$.

The Bayesian approach allows to base estimation of the component failure distributions
on both data and further knowledge not given by the data,
the latter usually provided in the form of expert knowledge.
This knowledge is encoded in form of a so-called prior distribution,
a distribution on the parameters of the component lifetime distributions.
This expert knowledge is especially important when there is very few data on the components (like in our scenario),
as only with its help meaningful estimates for the system reliability can be made.

@@

@@

The paper is organized as follows.
In Section~\ref{sec:weibull}, we describe a Bayesian analysis for Weibull component lifetimes
and illustrate the issue of prior-data conflict.
Section~\ref{sec:modforsurpr} then details the use of sets of priors
for the scale parameter $\lambda_k$ of the Weibull distribution,
showing how this mitigates the prior-data conflict issue for $\lambda_k$.

@@

System with components of $k=1,\ldots,K$ different types.
There are $n_k$ exchangeable components of type $k$ in the system.

For each $k$, component lifetimes ($i=1,\ldots,n_k$) are $T_i^k \mid \lambda_k \sim \wei(\kappa,\lambda_k)$,
where $\kappa$ is fixed and known:
\begin{linenomath*}
\begin{align}
f_k(t_i^k \mid \lambda_k) &= \frac{\kappa}{\lambda_k} (t_i^k)^{\kappa-1} e^{-\frac{(t_i^k)^{\kappa-1}}{\lambda_k}} \\
F_k(t_i^k \mid \lambda_k) &= 1 - e^{-\frac{(t_i^k)^\kappa}{\lambda_k}} = P_k(T_i^k \leq t_i^k \mid \lambda_k)
\end{align}
\end{linenomath*}

Conjugate prior is $\lambda_k \sim \ig(\alpha_k,\beta_k)$:
\begin{linenomath*}
\begin{align}
f_{\lambda_k}(\lambda_k\mid \alpha_k,\beta_k) &= \frac{(\beta_k)^{\alpha_k}}{\Gamma(\alpha_k)} \lambda_k^{-\alpha_k -1} e^{-\frac{\beta_k}{\lambda_k}}
\end{align}
\end{linenomath*}
In terms of canonical parameters $\nkz$ and $\ykz$, we assume $\lambda_k \mid \nkz,\ykz \sim \ig(\nkz + 1, \nkz\ykz)$.

>>>>

Often, there is no failure data for the system itself
(e.g., if the system is a prototype, or the system is used under unique circumstances),
but there is some information about failure times for the components the system is made of.
We are able to analyse systems of arbitrary system structures, %layout of components,
i.e., any combination and nesting of series, parallel, k-out-of-n, or bridge-type arrangements,
by use of the survival signature \cite{2012:survsign}.
In this paper, we assume that components can be divided into $K$ different groups,
and components within each group $k$ ($k=1, \ldots, K$) can be assumed to be exchangeable,
i.e., to follow the same failure time distribution.
We denote components of group $k$ as \emph{type $k$ components},
and assume independence between component types,
such that failure times of components of different types are independent.
We assume that type $k$ component lifetimes $T_i^k$ ($i = 1, \ldots, n_k$)
are Weibull distributed, with type-specific parameters $\beta_k$ (shape) and $\lambda_k$ (scale).
***type-specific shape not implemented yet!***

Focusing on the prototype application,
we assume that our observations consist solely of the failure times of components in this system up to time $\tnow$,
such that the failure times of components that have not failed by $\tnow$ are right-censored,
and calculate $\Rsys(t \mid t > \tnow) = P(\Tsys \geq t \mid t > \tnow)$,
which can be used to determine the remaining useful life of the system
(in short RUL, see, e.g., \citeNP{2014:rul-review}).
%This sparse information is complemented by expert information on component lifetimes $T_i^k$.

The Bayesian approach allows to base estimation of the component failure distributions
on both data and further knowledge not given by the data,
the latter usually provided in the form of expert knowledge.
This knowledge is encoded in form of a so-called prior distribution,
a distribution on the parameters of the component lifetime distributions.
Expert knowledge is especially important when there is very few data on the components (like in our scenario),
as only with its help meaningful estimates for the system reliability can be made.

@@

Due to the iterative nature of the Bayesian framework,
it is possible to use a set of posteriors based on component test data as the set of priors
instead of a purely expert-based set of priors as discussed so far.
In that case, prior-data conflict sensitivity allows to uncover also a conflict between
current observations in the running system and past observations from component tests.

@@

The paper is organized as follows.
In Section~\ref{sec:weibull}, we describe a Bayesian analysis for Weibull component lifetimes
and illustrate the issue of prior-data conflict.
Section~\ref{sec:modforsurpr} then details the use of sets of priors
for the scale parameter $\lambda_k$ of the Weibull distribution,
showing how this mitigates the prior-data conflict issue for $\lambda_k$.
***

@@

We consider a system with components of $k=1,\ldots,K$ different types;
for each type $k$, there are $n_k$ exchangeable components in the system.
For each type $k$ component, we assume for its lifetime $T_i^k$ ($i=1,\ldots,n_k$, $k = 1, \ldots, K$)
a Weibull distribution with fixed shape parameter $\beta_k > 0$, with density and cdf%
\footnote{Our approach would be possible also for other parametric lifetime distributions
that form a canonical exponential family
(see, e.g., \citeNP[p.~202 and 272f]{2000:bernardosmith}, or \citeNP[p.~8]{diss}).}
\begin{linenomath*}
\begin{align}
f_k(t_i^k \mid \lambda_k) &= \frac{\beta_k}{\lambda_k} (t_i^k)^{\beta_k-1} e^{-\frac{(t_i^k)^{\beta_k-1}}{\lambda_k}}\, \\
F_k(t_i^k \mid \lambda_k) &= 1 - e^{-\frac{(t_i^k)^\beta_k}{\lambda_k}} = P_k(T_i^k \leq t_i^k \mid \lambda_k)\,,
\end{align}
\end{linenomath*}
%For each $k$, component lifetimes ($i=1,\ldots,n_k$) are
in short $T_i^k \mid \lambda_k \sim \wei(\beta_k,\lambda_k)$.
%where $\beta_k$ is fixed and known:

Conjugate prior is $\lambda_k \sim \ig(a_k,b_k)$:
\begin{linenomath*}
\begin{align}
f_{\lambda_k}(\lambda_k\mid a_k,b_k) &= \frac{(b_k)^{a_k}}{\Gamma(a_k)} \lambda_k^{-a_k -1} e^{-\frac{b_k}{\lambda_k}}
\end{align}
\end{linenomath*}
In terms of canonical parameters $\nkz$ and $\ykz$, we assume $\lambda_k \mid \nkz,\ykz \sim \ig(\nkz + 1, \nkz\ykz)$.
